---
title: "Proteome Profiling"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

The following code is used to calculate and plot the ribosomal fraction of samples during exponential phase
```{r}

library(dplyr)
library(ggplot2)
library(cowplot)
library(tidyr)
library(stringr)
library(readr)

resDf_protein_trT_set00_StcYtcNasAgrNgrMgh_SYAN_baseMgAllMg_baseNaAllNa_ExpAllPhase_noMatchFilter_p1Sf_vst <- read_csv("~/Desktop/Wilke Lab/ecoli_multiple_growth_conditions-master/a_results/resDf_protein_trT_set00_StcYtcNasAgrNgrMgh_SYAN_baseMgAllMg_baseNaAllNa_ExpAllPhase_noMatchFilter_p1Sf_vst.csv")
 
nameDictionary_RNA_Protein <- read_csv("~/Desktop/nameDictionary_RNA_Protein.csv")

metaProtein <- read_csv("~/Desktop/Wilke Lab/ecoli_multiple_growth_conditions-master/a_results/metaProtein.csv")

#change the name of columns on data frames to enable joining of tables
names(nameDictionary_RNA_Protein)[names(nameDictionary_RNA_Protein) == 'b#'] <- 'b'
names(nameDictionary_RNA_Protein)[names(nameDictionary_RNA_Protein) == 'Protein_id'] <- 'gene_id'
names(resDf_protein_trT_set00_StcYtcNasAgrNgrMgh_SYAN_baseMgAllMg_baseNaAllNa_ExpAllPhase_noMatchFilter_p1Sf_vst)[names(resDf_protein_trT_set00_StcYtcNasAgrNgrMgh_SYAN_baseMgAllMg_baseNaAllNa_ExpAllPhase_noMatchFilter_p1Sf_vst) == 'X1'] <- 'gene_id'


#join normalized data to nameDictionary by gene_id
resDf_protein_trT_set00_StcYtcNasAgrNgrMgh_SYAN_baseMgAllMg_baseNaAllNa_ExpAllPhase_noMatchFilter_p1Sf_vst %>% left_join(nameDictionary_RNA_Protein, by = 'gene_id') %>% subset(select = -c(mRNA_ID, b)) %>% select(gene_id, gene_name, everything())-> df

df %>% subset(select = (-c(gene_id))) -> df

#sum gene counts for each media, filter only ribosomal genes, sum ribosomal gene counts for each media, calculate ribosomal fraction 
df %>% gather(media, count, MURI_016:MURI_140) %>% group_by(media) %>% mutate(total= sum(count)) %>% filter(str_detect(gene_name,regex('^rp'))) %>% group_by(media) %>% mutate(ribo = sum(count)) %>% mutate(fraction = ribo/total)-> total_ribo

total_ribo %>% subset(select = -c(gene_name,count, total, ribo )) %>% distinct(fraction) -> fraction 

#select growth condidtions and doubling times, and filter only exponential phase samples
metaProtein %>% select(growthPhase,carbonSource, dataSet,experiment,growthTime_hr,doublingTimeMinutes,Mg_mM,Na_mM,Mg_mM_Levels,Na_mM_Levels) %>% filter(growthPhase == 'exponential') -> df2

#calculate generations per hour
df2 %>% mutate(generations_per_hour = ((growthTime_hr*60)/doublingTimeMinutes)/growthTime_hr) -> df2

MyData <- read_csv("~/Desktop/MyData.csv", 
    col_types = cols(X1 = col_skip()))

#join Mydata (includes growth conditions and generation times) and fraction by sample
fraction %>% left_join(MyData, by = 'media') %>% na.omit() %>% select(-fraction,fraction)-> graphready

#average the ribosomal fractions of samples that have the same doubling time
graphready %>% group_by(doublingTimeMinutes) %>% mutate(fraction_avg = (sum(fraction))/n()) -> graphready

#unique fraction averages
graphready %>% distinct(fraction_avg, .keep_all = TRUE) -> final

#plot of fractions (non-averaged)
graphready %>% ggplot(aes(x=generations_per_hour, y=fraction, color=carbonSource)) + xlab('Generations per Hour') + ylab('Ribosomal Fraction') + geom_point() + geom_text(aes(label=media),hjust='inward', vjust=1 , size= 5)

#plot of unique fraction averages 
final %>% ggplot(aes(x=generations_per_hour, y=fraction_avg, color=carbonSource)) + xlab('Generations per Hour') + ylab('Ribosomal Fraction') + geom_point() + geom_text(aes(label=media),hjust='inward', vjust=1 , size= 5)

#get rid of NaCl and MgSO4 stress 
final_alt <- final[!grepl("stress", final$experiment),]
#get rid of NaCl and MgSO4 stress (non-unique fraction_avg)
final_alt2 <- graphready[!grepl("stress", graphready$experiment),]

final_alt2 %>% ggplot(aes(x=generations_per_hour, y=fraction, color=carbonSource)) + xlab('Generations per Hour') + ylab('Ribosomal Fraction') + geom_point() + geom_text(aes(label=media),hjust='inward', vjust=1 , size= 5) 

final_alt %>% ggplot(aes(x=generations_per_hour, y=fraction_avg, color=carbonSource)) + xlab('Generations per Hour') + ylab('Ribosomal Fraction') + geom_point() + geom_text(aes(label=media),hjust='inward', vjust=1 , size= 5) 

```

Na concentrations and NaCl stress
```{r}

#keep only ond samples on NaCL stress with unique Na concentrations (5,100,200,300)
graphready %>% filter(experiment == 'NaCl_stress') %>% ungroup() %>% distinct(Na_mM, carbonSource,.keep_all = TRUE) -> Na_final

graphready %>% ggplot(aes(x=generations_per_hour, y=fraction_avg, color=carbonSource)) + xlab('Generations per Hour') + ylab('Ribosomal Fraction') + geom_point() + geom_text(aes(label=Na_mM),hjust='inward', vjust=1 , size= 5) 

Na_final %>% ggplot(aes(x=generations_per_hour, y=fraction_avg, color=media)) + xlab('Generations per Hour') + ylab('Ribosomal Fraction') + geom_point() + geom_text(aes(label=Na_mM),hjust='inward', vjust=1 , size= 5) 

#non-distinct ribosomal fraction NaCl stress
graphready %>% filter(experiment == 'NaCl_stress') %>% ungroup() -> a

a %>% ggplot(aes(x=generations_per_hour, y=fraction, color=media)) + xlab('Generations per Hour') + ylab('Ribosomal Fraction') + geom_point() + geom_text(aes(label=Na_mM),hjust='inward', vjust=1 , size= 5) 

```

Mg concentrations and MgSO4 stress
``` {r}

#keep only ond samples on MgSO4 stress with unique Mg concentrations (0.08,0.8,8,200)
graphready %>% filter(experiment == 'MgSO4_stress_high') %>% ungroup() %>% distinct(Mg_mM, carbonSource,.keep_all = TRUE) -> Mg_final

graphready %>% ggplot(aes(x=generations_per_hour, y=fraction_avg, color=carbonSource)) + xlab('Generations per Hour') + ylab('Ribosomal Fraction') + geom_point() + geom_text(aes(label=Mg_mM),hjust='inward', vjust=1 , size= 5) 

Mg_final %>% ggplot(aes(x=generations_per_hour, y=fraction_avg, color=media)) + xlab('Generations per Hour') + ylab('Ribosomal Fraction') + geom_point() + geom_text(aes(label=Mg_mM),hjust='inward', vjust=1 , size= 5) 

#non-distinct ribosomal fraction MgSO4 stress
graphready %>% filter(experiment == 'MgSO4_stress_high') %>% ungroup() -> b

b %>% ggplot(aes(x=generations_per_hour, y=fraction, color=media)) + xlab('Generations per Hour') + ylab('Ribosomal Fraction') + geom_point() + geom_text(aes(label=Mg_mM),hjust='inward', vjust=1 , size= 5) 

```

The following code is used to calculate and plot the ribosomal fraction of samples during exponential phase using untransformed data

```{r}
Untransformed <- read_csv("~/Desktop/GitHub/proteome-profiling-ecoli/umut/resDf_protein_trT_set00_StcYtcNasAgrNgrMgh_SYAN_baseMgAllMg_baseNaAllNa_ExpAllPhase_noFilter_p1Sf_noNorm.csv")

names(Untransformed)[names(Untransformed) == 'X1'] <- 'gene_id'

#join untransformed data to nameDictionary by gene_id
Untransformed %>% left_join(nameDictionary_RNA_Protein, by = 'gene_id') %>% subset(select = -c(mRNA_ID, b)) %>% select(gene_name, everything())-> UTnames

UTnames %>% subset(select = (-c(gene_id))) -> UTnames

#sum gene counts for each media, filter only ribosomal genes, sum ribosomal gene counts for each media, calculate ribosomal fraction 
UTnames %>% gather(media, count, MURI_016:MURI_140) %>% group_by(media) %>% mutate(total= sum(count)) %>% filter(str_detect(gene_name,regex('^rp'))) %>% group_by(media) %>% mutate(ribo = sum(count)) %>% mutate(fraction = ribo/total)-> UTtotalribo

UTtotalribo %>% subset(select = -c(gene_name,count, total, ribo )) %>% distinct(fraction) -> UTfraction 

#join Mydata (includes growth conditions and generation times) and fraction by sample
UTfraction %>% left_join(MyData, by = 'media') %>% na.omit() %>% select(-fraction,fraction)-> UTready

#average the ribosomal fractions of samples that have the same doubling time
UTready %>% group_by(doublingTimeMinutes) %>% mutate(fraction_avg = (sum(fraction))/n()) -> UTready

#unique fraction averages
UTready %>% distinct(fraction_avg, .keep_all = TRUE) -> UTfinal

#plot of fractions
UTready %>% ggplot(aes(x=generations_per_hour, y=fraction, color=carbonSource)) + xlab('Generations per Hour') + ylab('Ribosomal Fraction') + geom_point() + geom_text(aes(label=media),hjust='inward', vjust=1 , size= 5)

#plot of fraction averages
UTfinal %>% ggplot(aes(x=generations_per_hour, y=fraction_avg, color=carbonSource)) + xlab('Generations per Hour') + ylab('Ribosomal Fraction') + geom_point() + geom_text(aes(label=media),hjust='inward', vjust=1 , size= 5)

#get rid of NaCl and MgSO4 stress 
UTfinal_alt <- UTfinal[!grepl("stress", UTfinal$experiment),]
#get rid of NaCl and MgSO4 stress (non-unique fraction_avg)
UTfinal_alt2 <- UTready[!grepl("stress", UTready$experiment),]

UTfinal_alt2 %>% ggplot(aes(x=generations_per_hour, y=fraction, color=carbonSource)) + xlab('Generations per Hour') + ylab('Ribosomal Fraction') + geom_point() + geom_text(aes(label=media),hjust='inward', vjust=1 , size= 5) 

UTfinal_alt %>% ggplot(aes(x=generations_per_hour, y=fraction_avg, color=carbonSource)) + xlab('Generations per Hour') + ylab('Ribosomal Fraction') + geom_point() + geom_text(aes(label=media),hjust='inward', vjust=1 , size= 5) 

```

Na concentrations and NaCl stress with untransformed data
``` {r}
#keep only ond samples on NaCL stress with unique Na concentrations (5,100,200,300)
UTready %>% filter(experiment == 'NaCl_stress') %>% ungroup() %>% distinct(Na_mM, carbonSource,.keep_all = TRUE) -> UTNa

UTNa %>% ggplot(aes(x=generations_per_hour, y=fraction_avg, color=media)) + xlab('Generations per Hour') + ylab('Ribosomal Fraction') + geom_point() + geom_text(aes(label=Na_mM),hjust='inward', vjust=1 , size= 5) 

#non-distinct ribosomal fraction NaCl stress
UTready %>% filter(experiment == 'NaCl_stress') %>% ungroup() -> UTa

UTa %>% ggplot(aes(x=generations_per_hour, y=fraction, color=media)) + xlab('Generations per Hour') + ylab('Ribosomal Fraction') + geom_point() + geom_text(aes(label=Na_mM),hjust='inward', vjust=1 , size= 5) 

```

Mg concentrations and MgSO4 stress
``` {r}
#keep only ond samples on MgSO4 stress with unique Mg concentrations (0.08,0.8,8,200)
UTready %>% filter(experiment == 'MgSO4_stress_high') %>% ungroup() %>% distinct(Mg_mM, carbonSource,.keep_all = TRUE) -> UTMg

UTMg %>% ggplot(aes(x=generations_per_hour, y=fraction_avg, color=media)) + xlab('Generations per Hour') + ylab('Ribosomal Fraction') + geom_point() + geom_text(aes(label=Mg_mM),hjust='inward', vjust=1 , size= 5) 

#non-distinct ribosomal fraction MgSO4 stress
UTready %>% filter(experiment == 'MgSO4_stress_high') %>% ungroup() -> UTb

UTb %>% ggplot(aes(x=generations_per_hour, y=fraction, color=media)) + xlab('Generations per Hour') + ylab('Ribosomal Fraction') + geom_point() + geom_text(aes(label=Mg_mM),hjust='inward', vjust=1 , size= 5) 

```
